Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-11 manuscript/manuscript.Rmd
models

# Fisheries: approximate models and approximate solutions

@@

*Total allowable catch (TAC)* is one such policy.
It consists on reducing the inclination of the line defined by $h(F)$ using a prefactor $\alpha$ in $h(F) = \alpha\times rF/2$. 
Plausible examples are $\alpha = 0.8$ or $0.9$.
Another common alternative is to have a prefactor $\alpha = \alpha(F)$ which decreases from one to zero as $F$ decreases below a threshold:
$$
  h(F) = \alpha(F)\times rF/2,
$$
where,
$$
  \alpha(F) = \begin{cases}
  1, \qquad &\text{if } F > F_{\text{thresh.}},\\
  F / F_{\text{thresh.}}, \qquad & \text{else}.
  \end{cases}
$$

>>>>

# Mathematical models of fisheries

In this section we introduce some models describing the population dynamics of fisheries.
In general, the class of models that appear in this context are *first order finite difference equations*.
For $n$ species, these models have the general form
\begin{align}
  \label{eq:general model}
  N_{t+1} - N_t = f(N_t) - h_t,
\end{align}
where $N(t) = (A_t,\ B_t,\ \dots) \in \mathbb{R}^{n}_+$ is a vector of populations, $f:\mathbb{R}^{n}\to\mathbb{R}^{n}$ is an arbitrary function, and $h_t$ is the harvest size at timestep $t$.

Most commonly, fisheries use 1-dimensional models, $n=1$.
One such model appearing in a variety of ecological contexts is *logistic growth*, for which
$$
  f(N_t) = r N_t\big(1 - N_t / K \big) =: L(N_t).
$$
More relevant to our current work, is the following single-species model studied in \cite{may77},
\begin{align}
  \label{eq:may}
  N_{t+1} - N_t
  = L(N_t) - \frac{\beta H N_t^2}{c^2 + N_t^2}.
\end{align}
The model has five parameters: a constant population $H$ of a species which preys on $N$, the maximal rate of predation $\beta$, and the predation half-maximum biomass $c$.

Part of what makes eq. \eqref{eq:may} an interesting study case is the presence of a *tipping point*.

# Fishery management approaches

@@

*Total allowable catch (TAC)* is one such policy.
It consists on reducing the inclination of the line defined by $h(F)$ using a prefactor $\alpha$ in $h(F) = \alpha\times rF/2$. 
Plausible examples are $\alpha = 0.8$ or $0.9$.
Another common alternative is to have a prefactor $\alpha = \alpha(F)$ which decreases from one to zero as $F$ decreases below a threshold:
$$
  h(F) = \alpha(F)\times rF/2,
$$
where,
$$
  \alpha(F) = \begin{cases}
  1, \qquad &\text{if } F > F_{\text{thresh.}},\\
  F / F_{\text{thresh.}}, \qquad & \text{else}.
  \end{cases}
$$
We call this policy *variable total allowable catch (VTAC)*.

## Dangers: Noisy parameters, noisy observations and model inaccuracy
